Robust Spectroscopic Inference with Imperfect Models

### 2014-12-05 tex/paper.tex
Open sourcing
Travis
trying requirements file
Kurucz fig updated
Partial comments addressed

\newcommand{\KK}{\mathcal{K}}
\newcommand{\Kglobal}{\KK^{\textrm{g}}}
\newcommand{\Klocal}{\KK^l}

@@

\shorttitle{Robust Spectroscopic Inference with Imperfect Models}
\shortauthors{Czekala et al.}

@@

\slugcomment{draft: \today{}}

@@

\author{Ian Czekala, Sean M.~Andrews, et al.}
\affil{Harvard-Smithsonian Center for Astrophysics, 60 Garden Street, Cambridge, MA 02138}
\email{iczekala@cfa.harvard.edu}

\begin{abstract}
We present a modular, extensible framework for the spectroscopic inference of 
physical parameters based on synthetic model spectra.  The subtraction of an 
imperfect model from a continuously sampled spectrum introduces covariance 
between adjacent datapoints (pixels) into the residual spectrum.  In the limit 
of high signal-to noise data with large spectral range that is common for 
stellar parameter estimation, that covariant structure can bias the parameter 
determinations.  We have designed a likelihood function formalism to account 
for the structure of the covariance matrix, utilizing the machinery of Gaussian 
process kernels.  We specifically address the common problem of mismatches in 
model spectral line strengths (with respect to data) due to intrinsic model 
imperfections (e.g., in the atomic or molecular data, or radiative transfer 
treatment) by developing a novel local covariance kernel framework that 
identifies and self-consistently downweights pathological spectral line 
``outliers."  By fitting multiple spectra in a hierarchical manner, these local 
kernels provide a mechanism to learn about and build data-driven corrections to 
synthetic model spectral libraries.  The application of this method, 
implemented as a freely available open source code, is demonstrated by fitting 
the high resolution optical ($V$-band) spectrum of WASP-14, an F5 dwarf with a 
transiting exoplanet, and the moderate resolution near-infrared ($K$-band) 
spectrum of Gliese~51, an M5 dwarf. 
\end{abstract}
\keywords{stars: fundamental parameters --- methods: data analysis --- methods: statistical}

@@

All astronomers recognize that spectroscopy offers a wealth of information that 
can help characterize the properties of the observing target.  In the context 
of stellar astrophysics, spectroscopy plays many fundamental roles.  The 
relative strengths and widths of stellar absorption lines provide access to key 
parameters like effective temperature ($T_{\rm eff}$) and surface gravity 
($\log g$), enabling model comparisons in the Hertzsprung-Russell diagram to 
estimate the masses and ages so crucial to understanding stellar evolution, as 
well as individual elemental (and molecular) abundances or the collective 
``metallicity" (typically parameterized as [Fe/H]), facilitating study of the 
chemical hallmarks of different stellar populations.  With sufficient spectral 
resolution, the velocity content of a spectrum can convey crucial information 
about stellar rotation ($v \sin i$) and kinematics (e.g., association with a 
cluster or companion through the radial velocity, $v_z$).  While many fields 
benefit from the spectroscopic measurements of these stellar properties, there 
is acute interest in the exoplanet community.  There, all estimates of the 
planet properties are made {\it relative} to the host properties (e.g., the 
planet-to-host mass or radius {\it ratio} is constrained with the radial 
velocity or transit techniques, respectively).  Moreover, essential clues to 
the planet formation process are encapsulated in the dependences of planet 
frequency on host mass \citep[e.g.,][]{johnson07,howard10} and metallicity 
\citep[e.g.,][]{fischer05,buchhave14}.

That said, the robust and quantitative extraction of physical (or empirical) parameters from an 
observed spectrum can be an extraordinary challenge.  Stellar models are ultimately required as a 
comparative benchmark to associate observed spectral features with the parameters of interest.  
Generating a synthetic model spectrum requires a complex numerical treatment of the stellar 
structure and radiative transfer through the atmosphere \citep[e.g.,][]{kurucz93,castelli04,
hauschildt99,husser13,paxton11}.  Detailed models calibrated to individual stars are important, but 
rare (e.g., the Sun, Vega); as such, these stellar models are relatively untested in large swaths 
of parameter-space.  Moreover, they necessarily include simplifications to treat complicated 
physical processes (e.g., convection) or computational limitations (e.g., geometry, boundary 
conditions), and often must rely on incomplete or inaccurate atomic and molecular information 
(e.g., oscillator strengths, opacities).  In principle, the models could be further improved with 
appropriate reference to spectroscopic datasets.  Nevertheless, they achieve a remarkable level of 
success in reproducing many key diagnostic features in stellar spectra.

There are various well-tested techniques being used to compare these models with observed spectra 
and thereby infer basic stellar parameters; here we highlight three examples to illustrate some key 
ideas.  First is a straightforward empirical approach that relies on distilling an information-rich 
subset of the data, usually in the form of spectral line equivalent widths and/or local continuum 
shapes.  A combined sequence of the ratios of these quantities are found to be especially sensitive 
to one key parameter in the models, and therefore can be used to quantify it \citep[e.g., {\tt 
MOOG};][]{sneden73,gray94,reid95,rojas-ayala10,rojas-ayala12}.  This ``indexing" approach has the 
advantage of being trivially fast, but each condensed relationship is only informative over a 
limited swath of parameter-space.  A second technique exploits the cross-correlation of a spectrum 
with a suite of model templates to optimize a set of parameters, usually with some weighting 
applied to specific spectral regions \citep[e.g., {\tt SPC};][]{buchhave12}.  In this case, the 
speed advantage is maintained (perhaps enhanced) and more data content is used (particularly in the 
spectral dimension), thereby achieving higher precision even for data with comparatively low 
sensitivity; the disadvantage is that the model quality and parameter inferences are assessed in an 
empirical, rather than probabilistic, framework.  A third approach employs a direct, pixel-by-pixel 
comparison between model and data, and is usually associated with a spectral synthesis back-end 
(e.g., {\tt SME}; \citealt{valenti96}).  This technique has the benefits of parametric flexibility 
(e.g., one can fit for arbitrary abundances or structures) and a proper inference framework 
\citep[usually a least-squares approach, although increasingly in a Bayesian format;][]{shkedy07,
schoenrich13}, but it is computationally expensive and can be biased by systematics 
\citep[e.g.,][]{mann13}; in practice, this often requires an effort focused on a pre-selected 
subsample of the data.

@@

The remainder of this Section describes the mechanics of this modeling framework.  First, a model 
spectrum is generated for a given set of physical parameters (Section \ref{subsec:synthetic}; 
Appendix~\ref{sec:Appendix}), and then post-processed to mimic reality using a set of observational 
and practical nuisance parameters (Section \ref{subsec:postprocess}).  Next, a direct, 
pixel-by-pixel comparison between the data and model spectra is made with a prescribed likelihood 
function and a parametric treatment of the covariances between pixel residuals (Section 
\ref{subsec:likelihood}).  That process is iterated under the guise of hierarchical Markov Chain 
Monte Carlo (MCMC) simulations to numerically explore the posterior probability density of the 
model conditioned on the data, and thereby to determine constraints on the parameters of interest 
(Section \ref{subsec:MCMC}).  Along the way, these procedures are illustrated with observations of 
the high resolution optical spectrum from a nearby F star.  That specific application, along with 
some alternative demonstrations of the method, are discussed in more detail in Section 
\ref{sec:examples}.

@@

We opt for the computationally expedient approach that employs a library of model spectra, 
$\flam(\{\vt_{\ast}\}^{\rm grid})$, where $\vt_{\ast} = [T_{\rm eff}, \,\, \log{g}, \,\, 
[{\rm Fe/H}]]$.  However, it is worth noting that the techniques developed here are applicable to 
{\it any} ``back-end" that generates a model spectrum.  In our adopted approach, the model spectrum 
for an arbitrary $\vt_{\ast}$ is interpolated from a spectral library,
\begin{equation} \label{eqn:interp} 
\flam(\{\vt_{\ast}\}^{\rm grid}) \leadsto \flam(\vt_{\ast}), 
\end{equation} 
where we assign the symbol $\leadsto$ as an interpolation operator.  The multi-dimensional 
interpolation in Eq.~\ref{eqn:interp} needs to be performed many times, so computational efficiency 
is critical.  In practice, a simple tri-linear interpolation is suitably fast, but introduces a 
non-negligible level of uncertainty.  These interpolation uncertainties can be empirically 
estimated by performing the operation in Eq.~\ref{eqn:interp} across a $\{\vt_{\ast}\}^{\rm grid}$ 
grid point, and then comparing the interpolated spectrum with the corresponding library spectrum 
\citep[see also][]{husser12}.  To suitably propagate the interpolation uncertainties into the 
likelihood calculations, we employ a Bayesian emulation technique that is described in detail in 
Appendix \ref{sec:Appendix}.

@@

Generally, the ``raw" model spectrum $\flam(\vt_{\ast})$ is highly over-sampled compared to the 
observed spectrum, and does not account for several additional observational and instrumental 
effects that become important in comparisons with real data.  Therefore, a certain amount of 
post-processing is required before assessing the model quality.  We treat that post-processing in 
two stages: the first deals with an additional set of ``observational" parameters, $\vt_{\rm obs}$, 
that incorporate dynamical effects, geometry, and the relative location of the target, while the 
second employs a suite of nuisance (hyper-)parameters ($\vp$) designed to mitigate an imperfect 
data calibration.

@@

% NEEDS UPDATED FIGURE
\begin{figure}[b!]
\includegraphics{figs/chebyshev.pdf}
\vspace{-0.75cm}
\figcaption{A demonstration of our treatment for residual calibration mismatches.  The observed 
spectra at the overlap of two echelle orders ({\it top}) have slightly ($\sim$1--3\%) discrepant 
shapes.  Chebyshev polynomials ({\it middle, top}) are designed to account for that mismatch ({\it 
middle, bottom}); rather, in practice, the model is scaled (multiplied) by those polynomials ({\it 
bottom}).  \label{fig:chebyshev}}
\end{figure}

@@

We employ an analogous, but more rigorous, approach to deal with this issue, cast {\it internal} to 
the modeling framework to propagate the uncertainties introduced by additional degrees of freedom 
in the model.  Deviations in the spectral shape introduced by calibration residuals are treated as 
explicit contributions to the model spectrum, enabling an exploration of the distribution of 
possible ``tweaks" to the calibration that can be compartmentalized into a set of nuisance 
parameters and then marginalized.  This way, the inferences for the stellar parameters will 
properly account for the underlying uncertainty in the calibration of the spectral shape.  In 
practice, this is achieved by distorting segments of the model spectrum with polynomials, 
$\mathsf{P}$ \citep[e.g.,][]{eisenstein06,koleva09}.  For data that has $N_{\rm ord}$ spectral 
orders, each denoted with index $o$, the model spectrum can be decomposed into the set of orders
\begin{eqnarray} \label{eqn:chebyshev}
\vM(\vT, \cheb) &=& \{ \vM_o(\vT) \times \mathsf{P}_o \} \\
                &=& \{ \vM_o(\vT) \times \sum_n c_o^{(n)} \, T_o^{(n)} \}, \nonumber
\end{eqnarray}
where $T^{(n)}$ is an $n^{\rm th}$ degree Chebyshev function.  The $n \times N_{\rm ord}$ 
coefficients are considered a set of nuisance (hyper-)parameters, $\cheb = [\{c_o^{(0)}, c_o^{(1)}, 
\ldots, c_o^{(n-1)} \}]$.  With judicious priors, we can ensure that the unintended treatment of 
real spectral features (e.g., molecular bands) as residual calibration artifacts is negligible.  
The lowest-degree (scaling) coefficient, $c^{(0)}$, is naturally degenerate with the solid angle, 
$\Omega$.  Therefore, we enforce an additional constraint that the mean of the polynomial is 
unity.  For data with a single spectral order, this means simply setting $c^{(0)} = 1$.  In the 
multiple order case, we assign $c^{(0)} = 1$ in an arbitrary order as an anchor, but permit the 
$c^{(0)}$ in other orders to be different.  Figure \ref{fig:chebyshev} offers a practical 
demonstration of how these nuisance parameters are applied to the model.

@@

The covariance matrix $\vC$ characterizes both the measurement uncertainty ($\sigma$; ``noise") in 
each pixel and the intrinsic covariance between pixels.  The special case where each pixel 
represents an independent measurement results in a diagonal covariance matrix, $\vC_{ij} = 
\delta_{ij} \,\, \sigma_i$ where $\sigma_i$ is the uncertainty in pixel $i$ and $\delta_{ij}$ is 
the Kronecker delta function, and Eq.~\ref{eqn:lnlikelihood} reduces to the familiar
\begin{equation}
\ln{p(\vD | \vM)} = -\frac{1}{2} \sum_i^{N_{\rm pix}} \frac{\vR_i^2}{\sigma_i^2} \equiv -\frac{\chi^2}{2},
\label{eqn:chisq}
\end{equation}
the sum of the square of the residuals weighted by their inverse variances.  However, the problem 
being addressed here necessitates the use of a more complex covariance matrix; additional 
off-diagonal terms that can explicitly characterize (1) pixel-to-pixel covariances imposed by the 
discrete over-sampling of the line-spread function, and (2) highly correlated residuals as 
manifestations of the still-imperfect model library are required to avoid biasing our inferences of 
the physically interesting parameters ($\vT$).  The following subsections describe how these issues 
are addressed in the practical implementation of $\vC$.

@@

\begin{figure}[!htb]
\begin{center}
  \includegraphics{residuals_23.pdf}
  \includegraphics{class0_residuals.pdf}
  \includegraphics{class0_autocorrelation.pdf}
  \figcaption{({\it top}) A comparison of the data and a typical model with parameters drawn from 
the posterior distribution, along with the corresponding residual spectrum.  ({\it middle}) A 
zoomed view of the gray band in the top panels, highlighting the mildly covariant residual 
structure that is produced by slight mismatches between the data and model spectra.  ({\it bottom}) 
The autocorrelation of the residual spectrum.  Notice the substantial autocorrelation signal for 
offsets of $\lesssim 8$ pixels, demonstrating clearly that the residuals are not described well by 
white (Poisson) noise alone. \label{fig:class0}}
\end{center}
\end{figure}

@@

For a well-designed spectrograph and sufficiently accurate model, this {\it global} (i.e., present 
throughout the spectrum) covariant structure should have a relatively low amplitude and small 
correlation length.  To describe that structure, we use a stationary covariance kernel (or radial 
basis function) with an amplitude that depends only on the velocity distance between two pixels, 
\begin{equation}
  r_{ij} = r(\lambda_i, \lambda_j) = \Delta v = \frac{c}{2} \left | \frac{\lambda_i 
   - \lambda_j}{ \lambda_i + \lambda_j} \right |,
\end{equation}
where $c$ is the speed of light.  This parametric kernel describes the covariance between pixel 
residuals, 
\begin{equation}
  \Kglobal_{ij} =  \langle \vR_i \; \vR_j \rangle.
  \label{eqn:expectation}
\end{equation}
A variety of kernels have been used in the field of Gaussian processes to parameterize such a 
covariant structure \citep[e.g.,][]{rasmussen05}.  Here, we adopt the commonly-employed \matern\ 
kernel with $\nu = 3/2$,
\begin{equation}
  \Kglobal_{ij}(\vp_{{\mathsf C}, {\rm g}}) = a_{\rm g} \left(1 + \frac{\sqrt{3}\, r_{ij}}{\ell} \right ) \exp 
   \left (- \frac{\sqrt{3}\, r_{ij}}{\ell} \right ),
   \label{eqn:global}
\end{equation}
where the (hyper-)parameters $\vp_{{\mathsf C}, {\rm g}} = [a_{\rm g}, \ell]$ include an amplitude 
($a_{\rm g}$) and scale ($\ell$).  To ensure that $\vC$ remains a relatively sparse matrix that 
enables computational expendiency, we employ a Hann window function
\begin{equation}
  w_{ij}^{\rm g} \,(r_0) = \left \{ 
    \begin{array}{cc}
    \frac{1}{2} + \frac{1}{2} \cos \left(\frac{\pi r_{ij}}{r_0} \right) & r_{ij} \le r_0 \\
    0 & r_{ij} > r_0 \\
  \end{array}
  \right .
  \label{eqn:Hann}
\end{equation}
as a taper.  The truncation distance $r_0$ can be fixed to a reasonable multiple of the scale (we 
set $r_0 = 4\ell$).

@@

\begin{figure*}[!htb]
\begin{center}
\includegraphics[scale=1.05]{figs/matrix_compilation.pdf}
\figcaption{A decomposition of the modeling procedure, explicitly highlighting the roles of the 
various contributions to the covariance matrix.  The top panels show a typical comparison between 
the data and model spectrum, along with the associated residual spectrum.  The subsequent panels 
focus on the illustrative region shaded in grey.  The left column of panels show the corresponding 
region of the covariance matrix $\vC$, decomposed into its primary contributions: from top to 
bottom, the trivial noise matrix, then combined with the global covariance kernel, and finally 
including an appropriate local covariance kernel.  In the right column, we show the zoomed-in 
residual spectrum ({\it black}) along with example random draws from the subsets of $\vC$ exhibited 
to the left.  The shaded contours ({\it orange}) represent the 1, 2, and 3\,$\sigma$ dispersions of 
an ensemble of 200 random draws from $\vC$.  Note that the trivial noise matrix ($\delta_{ij} 
\sigma_i$) poorly reproduces both the scale and structure of the residual spectrum.  The addition 
of a global kernel ($\mathcal{K}^{\rm g}$) more closely approximates the structure and amplitude of 
the residuals, but misses the outlier line at 5202.2\,\AA.  Including a local kernel 
($\mathcal{K}^l$) at that location results in a covariance structure that does an excellent job 
reproducing all the key residual features.  \label{fig:matrix}}
\end{center}
\end{figure*}

>>>>

\newcommand{\KK}{\mathcal{K}}
\newcommand{\Kglobal}{\KK^{\textrm{G}}}
\newcommand{\Klocal}{\KK^{\textrm{L}}}

@@

\shorttitle{Robust Spectroscopic Inference}
\shortauthors{Czekala et al.}

@@

@@

\newcommand{\harvard}{1}
\newcommand{\nyu}{2}
\newcommand{\mpia}{3}
\newcommand{\cds}{4}

\author{%
  Ian~Czekala\altaffilmark{\harvard }, 
  Sean~M.~Andrews\altaffilmark{\harvard },
  Kaisey~S.~Mandel\altaffilmark{\harvard },
  David~W.~Hogg\altaffilmark{\nyu },
  and Gregory~M.~Green\altaffilmark{\harvard}
}

\email{iczekala@cfa.harvard.edu}
\altaffiltext{\harvard}  {Harvard-Smithsonian Center for Astrophysics, 
			 60 Garden Street, Cambridge, MA 02138}
\altaffiltext{\nyu}      {Center for Cosmology and Particle Physics,
                          Department of Physics, New York University,
                          4 Washington Place, New York, NY, 10003, USA}
%\altaffiltext{\mpia}     {Max-Planck-Institut f\"ur Astronomie,
%			  K\"onigstuhl 17, D-69117 Heidelberg, Germany}
%\altaffiltext{\cds}      {Center for Data Science,
%                          New York University,
%                          4 Washington Place, New York, NY, 10003, USA}

\begin{abstract}
We present a modular, extensible framework for the spectroscopic inference of 
physical parameters based on synthetic model spectra.  The subtraction of an 
imperfect model from a continuously sampled spectrum introduces covariance 
between adjacent datapoints (pixels) into the residual spectrum.  In the limit 
of high signal-to-noise data with large spectral range that is common for 
stellar parameter estimation, that covariant structure can bias the parameter 
determinations.  We have designed a likelihood function formalism to account 
for the structure of the covariance matrix, utilizing the machinery of Gaussian 
process kernels.  We specifically address the common problem of mismatches in 
model spectral line strengths (with respect to data) due to intrinsic model 
imperfections (e.g., in the atomic or molecular data, or radiative transfer 
treatment) by developing a novel local covariance kernel framework that 
identifies and self-consistently downweights pathological spectral line 
``outliers."  By fitting multiple spectra in a hierarchical manner, these local 
kernels provide a mechanism to learn about and build data-driven corrections to 
synthetic model spectral libraries.  The application of this method, 
implemented as a freely available open source code, is demonstrated by fitting 
the high resolution optical ($V$-band) spectrum of WASP-14, an F5 dwarf with a 
transiting exoplanet, and the moderate resolution near-infrared ($K$-band) 
spectrum of Gliese~51, an M5 dwarf. 
\end{abstract}
\keywords{stars: fundamental parameters --- techniques: spectroscopic --- stars: late-type --- 
  stars: statistics --- methods: data analysis --- methods: statistical}

@@

All astronomers recognize that spectroscopy offers a wealth of information that can help
characterize the properties of the observing target.  In the context of stellar astrophysics,
spectroscopy plays many fundamental roles.  The relative strengths and widths of stellar absorption
lines provide access to key parameters like effective temperature ($T_{\rm eff}$) and surface
gravity ($\log g$), enabling model comparisons in the Hertzsprung-Russell diagram to estimate the
masses and ages so crucial to understanding stellar evolution, as well as individual elemental (and
molecular) abundances or the collective ``metallicity" (typically parameterized as [Fe/H]),
facilitating study of the chemical hallmarks of different stellar populations.  With sufficient
spectral resolution, the velocity content of a spectrum can convey crucial information about stellar
rotation ($v \sin i$) and kinematics (e.g., association with a cluster or companion through the
radial velocity, $v_z$).  While many fields benefit from the spectroscopic measurements of these
stellar properties, they are of acute interest to the exoplanet community.  There, all estimates of
the planet properties are made {\it relative} to the host properties (e.g., the planet-to-host mass
function or radius {\it ratio} is constrained with the radial velocity or transit techniques,
respectively).  Moreover, essential clues to the planet formation process are encapsulated in the
dependences of planet frequency on host mass \citep[e.g.,][]{johnson07,howard10} and metallicity
\citep[e.g.,][]{fischer05,buchhave14}.

The robust and quantitative extraction of physical (or empirical) parameters from an 
observed spectrum can be an extraordinary challenge.  Stellar models are ultimately required as a 
comparative benchmark to associate observed spectral features with the parameters of interest.  
Generating a synthetic model spectrum requires a complex numerical treatment of the stellar 
structure and radiative transfer through the atmosphere \citep[e.g.,][]{kurucz93,castelli04,
hauschildt99,husser13,paxton11}.  Detailed models calibrated to individual stars are important, but 
rare (e.g., the Sun, Vega); as such, these stellar models are relatively untested in large swaths 
of parameter-space.  Moreover, they necessarily include simplifications to treat complicated 
physical processes (e.g., convection) or computational limitations (e.g., geometry, boundary 
conditions), and often must rely on incomplete or inaccurate atomic and molecular information 
(e.g., oscillator strengths, opacities).  In principle, the models could be further improved with 
appropriate reference to spectroscopic datasets.  Nevertheless, they achieve a remarkable level of 
success in reproducing many key diagnostic features in stellar spectra.

There are various well-tested techniques being used to compare these models with observed spectra
and thereby infer basic stellar parameters; here we highlight three examples to illustrate some key
ideas.  First is a straightforward empirical approach that relies on distilling an information-rich
subset of the data, usually in the form of spectral line equivalent widths and/or local continuum
shapes.  A combined sequence of the ratios of these quantities are found to be especially sensitive
to one key parameter in the models, and therefore can be used to quantify it \citep[e.g., {\tt
MOOG};][]{sneden73,gray94,reid95,rojas-ayala10,rojas-ayala12}.  This ``indexing" approach has the
advantage of being trivially fast, but each condensed relationship is only informative over a
limited swath of parameter-space.  A second technique exploits the cross-correlation of a spectrum
with a suite of model templates to optimize a set of parameters, usually with some weighting applied
to specific spectral regions \citep[e.g., {\tt SPC};][]{buchhave12}.  In this case, the speed
advantage is maintained (perhaps enhanced) and more data content is used (particularly in the
spectral dimension), thereby achieving higher precision even for data with comparatively low
signal-to-noise; the disadvantage is that the model quality and parameter inferences are assessed in
an heuristic, rather than probabilistic, framework, making it difficult to quantify uncertainty in
the stellar parameters without a well-defined model of how the data is generated.  A third approach
employs a direct, pixel-by-pixel comparison between model and data, and is usually associated with a
spectral synthesis back-end (e.g., {\tt SME}; \citealt{valenti96}).  This technique has the benefits
of parametric flexibility (e.g., one can fit for arbitrary abundances or structures) and a proper
inference framework \citep[usually a least-squares approach, although increasingly in a Bayesian
format;][]{shkedy07, schoenrich13}, but it is computationally expensive and can be biased by
systematics \citep[e.g.,][]{mann13}; in practice, this often requires an effort focused on a
pre-selected subsample of the data.

@@

The remainder of this Section describes the mechanics of this modeling framework.  First, a model
spectrum is generated for a given set of physical parameters (Section \ref{subsec:synthetic};
Appendix~\ref{sec:Appendix}), and then post-processed to mimic reality using a set of observational
and practical nuisance parameters (Section \ref{subsec:postprocess}).  Next, a direct,
pixel-by-pixel comparison between the data and model spectra is made with a prescribed likelihood
function and a parametric treatment of the covariances between pixel residuals (Section
\ref{subsec:likelihood}).  That process is iterated using Markov Chain Monte Carlo (MCMC)
simulations in a multi-stage Gibbs sampler to numerically explore the posterior probability density
of the model conditioned on the data, and thereby to determine constraints on the parameters of
interest (Section \ref{subsec:MCMC}).  Along the way, these procedures are illustrated with
observations of the high resolution optical spectrum from a nearby F star.  That specific
application, along with some alternative demonstrations of the method, are discussed in more detail
in Section \ref{sec:examples}.

@@

We opt for the computationally expedient approach that employs a library of model spectra, 
$\flam(\{\vt_{\ast}\}^{\rm grid})$, where $\vt_{\ast} = [T_{\rm eff}, \,\, \log{g}, \,\, 
[{\rm Fe/H}]]$.  However, it is worth noting that the techniques developed here are applicable to 
{\it any} ``back-end" that generates a model spectrum.  In our adopted approach, the model spectrum 
for an arbitrary $\vt_{\ast}$ is interpolated from a spectral library,
\begin{equation} \label{eqn:interp} 
\flam(\{\vt_{\ast}\}^{\rm grid}) \leadsto \flam(\vt_{\ast}), 
\end{equation} 
where we assign the symbol $\leadsto$ as an interpolation operator.  The multi-dimensional
interpolation in Eq.~\ref{eqn:interp} needs to be performed many times, so computational efficiency
is critical.  In practice, a simple linear interpolation from nearby grid points in the spectral
library is suitably fast, but introduces a non-negligible level of interpolation error. Although
generally small, this error can significantly bias the inference of stellar parameters
when the signal-to-noise of the data is sufficiently high. To account for the uncertainty in a given
interpolated spectrum, we design a Bayesian spectral emulator, which we describe in detail in
Appendix \ref{sec:Appendix}. Rather than delivering a single interpolated spectrum for a given
$\vt_\ast$, the emulator delivers a probability distribution over a set of possible interpolated
spectra. By marginalizing over all of the possible interpolated models, we can modify the likelihood
function (introduced in Section~\ref{subsec:likelihood}) to account for the interpolation uncertainty. For
simplicity, however, we now proceed as if we used linear interpolation from a synthetic library to
generate model spectra and that the error introduced by this process is negligible, since the
following concepts are applicable regardless of the interpolation method.

@@

Generally, the ``raw" model spectrum $\flam(\vt_{\ast})$ is highly over-sampled compared to the 
observed spectrum, and does not account for several additional observational and instrumental 
effects that become important in comparisons with real data.  Therefore, a certain amount of 
post-processing is required before assessing the model quality.  We treat that post-processing in 
two stages: the first deals with an additional set of ``observational" parameters, $\vt_{\rm obs}$, 
that incorporate dynamical effects, geometry, and the relative location of the target, while the 
second employs a suite of nuisance parameters ($\vp$) designed to mitigate an imperfect 
data calibration.

@@

\begin{figure}[b!]
\includegraphics{figs/chebyshev.pdf}
\vspace{-0.75cm}
\figcaption{A demonstration of our treatment for residual calibration mismatches.  The observed 
spectra at the overlap of two echelle orders ({\it top}) have slightly ($\sim$1--3\%) discrepant 
shapes.  By using Chebyshev polynomials ({\it middle, top}) one can correct for that mismatch by
adjusting the data ({\it middle, bottom}); instead, in practice we equivalently distort the model
by these polynomials ({\it bottom}) such that the model remains linear in the Chebyshev coefficients
(Eq.~\ref{eqn:chebyshev}).  \label{fig:chebyshev}}
\end{figure}

@@

We employ an analogous, but more rigorous, approach to deal with this issue, cast {\it internally}
to the modeling framework to propagate the uncertainties introduced by additional degrees of freedom
in the model.  Deviations in the spectral shape introduced by calibration residuals are treated as
explicit contributions to the model spectrum, enabling an exploration of the distribution of
possible ``tweaks" to the calibration that can be compartmentalized into a set of nuisance
parameters and then marginalized.  This way, the inferences for the stellar parameters will properly
account for the underlying uncertainty in the calibration of the spectral shape.  In practice, this
is achieved by distorting segments of the model spectrum with polynomials, $\mathsf{P}$
\citep[e.g.,][]{eisenstein06,koleva09}. Figure \ref{fig:chebyshev} offers a practical demonstration
of how these nuisance parameters are applied to the model. For data that has $N_{\rm ord}$ spectral
orders, each denoted with index $o$, the model spectrum can be decomposed into the set of orders
\begin{eqnarray} \label{eqn:chebyshev}
\vM(\vT, \cheb) &=& \{ \vM_o(\vT) \times \mathsf{P}_o \} \\
                &=& \{ \vM_o(\vT) \times \sum_n c_o^{(n)} \, T_o^{(n)} \}, \nonumber
\end{eqnarray}
where $T^{(n)}$ is an $n^{\rm th}$ degree Chebyshev function.  The $n \times N_{\rm ord}$ 
coefficients are considered a set of nuisance parameters, $\cheb = [\{c_o^{(0)}, c_o^{(1)}, 
\ldots, c_o^{(n-1)} \}]$.  With judicious priors, we can ensure that the unintended treatment of 
real spectral features (e.g., molecular bands) as residual calibration artifacts is negligible.  
The lowest-degree (scaling) coefficient, $c^{(0)}$, is naturally degenerate with the solid angle, 
$\Omega$.  Therefore, we enforce an additional constraint that the mean of the polynomial is 
unity.  For data with a single spectral order, this means simply setting $c^{(0)} = 1$.  In the 
multiple order case, we assign $c^{(0)} = 1$ in an arbitrary order as an anchor, but permit the 
$c^{(0)}$ in other orders to be different.

@@

The covariance matrix $\vC$ characterizes both the measurement uncertainty ($\sigma$; ``noise") in 
each pixel and the intrinsic covariance between pixels.  The special case where each pixel 
represents an independent measurement results in a diagonal covariance matrix, $\vC_{ij} = 
\delta_{ij} \,\, \sigma_i$ where $\sigma_i$ is the uncertainty in pixel $i$ and $\delta_{ij}$ is 
the Kronecker delta function. Under the assumption that the covariance matrix is held constant, Eq.~\ref{eqn:lnlikelihood} reduces to the familiar
\begin{equation}
  \ln{p(\vD | \vM)} - \textrm{constant} = -\frac{1}{2} \sum_i^{N_{\rm pix}} \frac{\vR_i^2}{\sigma_i^2} \equiv -\frac{\chi^2}{2},
\label{eqn:chisq}
\end{equation}
the sum of the square of the residuals weighted by their inverse variances.  However, the problem 
being addressed here necessitates the use of a more complex covariance matrix; additional 
off-diagonal terms that can explicitly characterize (1) pixel-to-pixel covariances imposed by the 
discrete over-sampling of the line-spread function, and (2) highly correlated residuals as 
manifestations of the still-imperfect model library are required to avoid biasing our inferences of 
the physically interesting parameters ($\vT$).  The following subsections describe how these issues 
are addressed in the practical implementation of $\vC$.

@@

\begin{figure}[!htb]
\begin{center}
  \includegraphics{residuals_23.pdf}
  \includegraphics{class0_residuals.pdf}
  \includegraphics{class0_autocorrelation.pdf}
  \figcaption{({\it top}) A comparison of the data and a typical model with parameters drawn from 
the posterior distribution, along with the corresponding residual spectrum.  ({\it middle}) A 
zoomed view of the gray band in the top panels, highlighting the mildly covariant residual 
structure that is produced by slight mismatches between the data and model spectra.  ({\it bottom}) 
The autocorrelation of the residual spectrum.  Notice the substantial autocorrelation signal for 
offsets of $\lesssim 8$ pixels, demonstrating clearly that the residuals are not well described by 
white (Poisson) noise alone. \label{fig:class0}}
\end{center}
\end{figure}

@@

For a well-designed spectrograph and sufficiently accurate model, this {\it global} (i.e., present 
throughout the spectrum) covariance should have a relatively low amplitude and small 
correlation length.  To describe that structure, we use a stationary covariance kernel (or radial 
basis function) with an amplitude that depends only on the velocity distance between two pixels, 
\begin{equation}
  r_{ij} = r(\lambda_i, \lambda_j) = \Delta v = \frac{c}{2} \left | \frac{\lambda_i 
   - \lambda_j}{ \lambda_i + \lambda_j} \right |,
\end{equation}
where $c$ is the speed of light.  This parametric kernel describes the covariance between pixel 
residuals, 
\begin{equation}
  \Kglobal_{ij} =  \langle \vR_i \; \vR_j \rangle.
  \label{eqn:expectation}
\end{equation}
A variety of kernels have been used in the field of Gaussian processes to parameterize such a 
covariant structure \citep[e.g.,][]{rasmussen05}.  Here, we adopt the commonly-employed \matern\ 
kernel with $\nu = 3/2$,
\begin{equation}
  \Kglobal_{ij}(\vp_{{\mathsf C}, {\rm g}}) = a_{\rm g} \left(1 + \frac{\sqrt{3}\, r_{ij}}{\ell} \right ) \exp 
   \left (- \frac{\sqrt{3}\, r_{ij}}{\ell} \right ),
   \label{eqn:global}
\end{equation}
where the hyperparameters $\vp_{{\mathsf C}, {\rm g}} = [a_{\rm g}, \ell]$ include an amplitude 
($a_{\rm g}$) and scale ($\ell$).  To ensure that $\vC$ remains a relatively sparse matrix that 
enables computational expediency, we employ a Hann window function
\begin{equation}
  w_{ij}^{\rm G} \,(r_0) = \left \{ 
    \begin{array}{cc}
    \frac{1}{2} + \frac{1}{2} \cos \left(\frac{\pi r_{ij}}{r_0} \right) & r_{ij} \le r_0 \\
    0 & r_{ij} > r_0 \\
  \end{array}
  \right .
  \label{eqn:Hann}
\end{equation}
as a taper.  The truncation distance $r_0$ can be fixed to a reasonable multiple of the scale (we 
set $r_0 = 4\ell$).

@@

\begin{figure*}[!htb]
\begin{center}
\includegraphics[scale=1.05]{figs/matrix_compilation.pdf}
\figcaption{A decomposition of the modeling procedure, explicitly highlighting the roles of the 
various contributions to the covariance matrix.  The top panels show a typical comparison between 
the data and model spectrum, along with the associated residual spectrum.  The subsequent panels 
focus on the illustrative region shaded in grey.  The left column of panels show the corresponding 
region of the covariance matrix $\vC$, decomposed into its primary contributions: from top to 
bottom, the trivial noise matrix, then combined with the global covariance kernel, and finally 
including an appropriate local covariance kernel.  In the right column, we show the zoomed-in 
residual spectrum ({\it black}) along with example random draws from the subsets of $\vC$ exhibited 
to the left.  The shaded contours ({\it orange}) represent the 1, 2, and 3\,$\sigma$ dispersions of 
an ensemble of 200 random draws from $\vC$.  Note that the trivial noise matrix ($\delta_{ij} 
\sigma_i$) poorly reproduces both the scale and structure of the residual spectrum.  The addition 
of a global kernel ($\mathcal{K}^\textrm{G}$) more closely approximates the structure and amplitude of 
the residuals, but misses the outlier line at 5202.2\,\AA.  Including a local kernel 
($\mathcal{K}^\textrm{L}$) at that location results in a covariance structure that does an excellent job 
reproducing all the key residual features.  \label{fig:matrix}}
\end{center}
\end{figure*}
